Unsupervised embeddings of organizations capture latent structure of scientific mobility

### 2020-04-20 paper/current/main.tex
typo
Minor changes while skimming main-text

When scientists move, they bring their knowledge with them, and so understanding the flows of scientists allows us to understand the flow of knowledge around the world. 
However, given its complexity~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics}, new approaches are needed to make sense of landscape of global scientific mobility. 
A basic component of mobility is distance--people are more likely to move nearby than to cross continents.
Although many alternative ways to quantify the distance~\autocite{boschma2005proximity} exist, geographic distance remains an unambiguous and foundational quantity for the distance.
In turn, models of human mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation models ~\autocite{simini2012universal}, rely almost exclusively on geographic distance.
However, distance is not always so straightforward as simple geography. 
Geographic distance fails to capture the host of historical, linguistic, cultural, and cognitive distance that also drive mobility. 
For instance, Australia may be physically close to Indonesia, but in terms of language, history, and shared culture, they are essentially neighbors to the United Kingdom.
Then, how can we quantify meaningful, useful distance between institutions, cities, and countries?

Here, we propose a neural embedding model as an efficient means of encoding an effective measure of distance using large-scale data on mobility trajectories.
It has been demonstrated that the neural embedding model preserves semantic relationships between words in a dense, continuous and meaningful low-dimensional vector representation~\autocite{mikolov2013word2vec}, and has been applied to reveal insights in text~\autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}.
In addition to encoding latent structure, this approach is computationally efficient, scalable for large datasets, robust to noise, and allows for a range of unique analyses that leverage their semantic properties using simple vector arithmetic.

@@

%
% Figure - Example of mobility sentences
%
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{\figdir/mobility_example_combined.pdf}
    \caption{
        \textbf{Co-affiliation trajectories represented as sentences}
        \textbf{a.}
        An author published five papers across five time periods, with only one affiliation listed in the byline of each paper.
        To each organization, a unique identifier is assigned, and they are assembled into a mobility sentence.
        \textbf{b.}
        In a more complex case, an author lists several affiliations on the byline of papers in $t_{2}$, $t_{3}$, and $t_{4}$. Here, organization ids are still ordered by year, however order within a single year is random.
    }
    \label{fig:methods:mobility_sentence}
\end{figure}

@@

To rewrite  Eq.\ref{eq:gravity_model} in terms of $T_{ij}$, we replace  $M_*$ with $\mu_{i*}$, the total flux to or from $i$ which can be written as

@@

where $N(i)\text{, } N(j) \text{ and } N(i,j)$ denotes the number of times word $i$, $j$, and word-context-pair ($i$, $j$) in the corpus respectively. Variable $k$ is the number of negative samples and $|D|$ is the size of the corpus.

From Matrix Eq.\ref{eq:sgns_relation}, we can get embedding vector with rank-d approximation by SVD. The rank-d SVD factorize matrix A into product of three matrices  $L_d \Sigma R_d^T$, where $L_d$ and $R_d$ is $n \times d$ real matrices, and $\Sigma$ is    $d \times d$ diagonal matrix.

@@

Scientific mobility is complex and can be defined in many possible ways~\autocite{robinson2019mobility}.
Here, authors are classified as mobile when they have at least two distinct organization ids in their trajectory, meaning that they have published using two or more distinct affiliations between 2008 and 2019.
Under this definition, mobile authors constitute 3,007,192 or 23.2 percent of all authors, and 17,700,095 author-affiliation combinations.
Mobile authors were associated with, on average, 2.5 distinct organizational affiliations.
Rates of mobility tended to differ across country.
For example, France, Qatar, the USA, Iraq, and Luxembourg had the most mobile authors (Fig.~\ref{fig:supp:descriptives}c).
However, due to their size, the USA, accounted for nearly 40 percent of all mobile authors worldwide (Fig.~\ref{fig:supp:descriptives}a), with 10 countries accounting for 80 percent of all mobility (Fig.~\ref{fig:supp:descriptives}b).
In most cases, countries with a high degree inter-organization mobility also had a high degree of international mobility;
however, the USA and France had a high degree of inter-organization mobility with the corresponding international mobility (Fig.~\ref{fig:supp:descriptives}c).
While the number of publications has increased year-to-year, the mobility and disciplinary makeup of the dataset has not significantly changed across the period of study (Fig.~\ref{fig:supp:pubs_over_time}).

>>>>

When scientists move, they bring their knowledge with them, and so understanding the flows of scientists allows us to understand the flow of knowledge around the world. 
However, given its complexity~\autocite{robinson2019mobility, vannoorden2012mobility, deville2014career, ackers2008mobility, scott2015dynamics}, new approaches are needed to make sense of landscape of global scientific mobility. 
A basic component of mobility is distance--people are more likely to move nearby than to cross continents.
Although many alternative ways to quantify the distance exist~\autocite{boschma2005proximity}, geographic distance remains the unambiguous and foundational quantity.
In turn, models of human mobility, such as the Gravity~\autocite{zipf1946gravity} and Radiation models ~\autocite{simini2012universal}, rely almost exclusively on geographic distance.
However, distance is not always so straightforward as simple geography. 
Geographic distance alone fails to capture the host of historical, linguistic, cultural, and cognitive distances that also drive mobility. 
For instance, Australia may be physically close to Indonesia, but in terms of language, history, and shared culture, they are essentially neighbors to the United Kingdom.
Then, how can we quantify meaningful, useful distance between institutions, cities, and countries?

It has been demonstrated that the neural embedding model preserves semantic relationships between words in a dense, continuous and meaningful low-dimensional vector representation~\autocite{mikolov2013word2vec}, and has been applied to reveal insights in text~\autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}.
In addition to encoding latent structure, this approach is computationally efficient, scalable for large datasets, robust to noise, and allows for a range of unique analyses that leverage their semantic properties using simple vector arithmetic.

@@

%
% Figure - Example of mobility sentences
%
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{\figdir/mobility_example_combined.pdf}
    \caption{
        \textbf{Co-affiliation trajectories represented as sentences}
        \textbf{a.}
        An author published five papers across five time periods, with only one affiliation listed in the byline of each paper.
        To each organization, a unique identifier is assigned, and they are assembled into a mobility sentence ordered by year of publication.
        \textbf{b.}
        In a more complex case, an author lists several affiliations on the byline of papers in $t_{2}$, $t_{3}$, and $t_{4}$. 
        In the case where multiple organizations are listed within the same year, then organization IDs within a single year are randomized across each training iteration or "epoch" of the \textit{word2vec} model. 
    }
    \label{fig:methods:mobility_sentence}
\end{figure}

@@

To rewrite  Eq.\ref{eq:gravity_model} in terms of $T_{ij}$, we replace  $M_i$ with $\mu_{i}$, the total flux to or from $i$ which can be written as

@@

where $N(i)\text{, } N(j) \text{ and } N(i,j)$ denotes the number of times word $i$, $j$, and word-context-pair ($i$, $j$) in the corpus respectively. Variable $k$ is the number of negative samples and $|D|$ is the size of the corpus. From Matrix Eq.\ref{eq:sgns_relation}, we can get embedding vector with rank-d approximation by SVD. The rank-d SVD factorize matrix A into product of three matrices  $L_d \Sigma R_d^T$, where $L_d$ and $R_d$ is $n \times d$ real matrices, and $\Sigma$ is    $d \times d$ diagonal matrix.

@@

Scientific mobility is complex and can be defined in many possible ways~\autocite{robinson2019mobility} and at multiple levels of analysis.
Here, authors are classified as mobile when they have at least two distinct organization ids in their trajectory, meaning that they have published using two or more distinct affiliations between 2008 and 2019.
Under this definition, mobile authors constitute 3,007,192 or 23.2 percent of all authors, and 17,700,095 author-affiliation combinations.
Mobile authors were associated with, on average, 2.5 distinct organizational affiliations.
Rates of mobility tended to differ across country.
For example, France, Qatar, the USA, Iraq, and Luxembourg had the most mobile authors (Fig.~\ref{fig:supp:descriptives}c).
However, due to their size, the USA, accounted for nearly 40 percent of all mobile authors worldwide (Fig.~\ref{fig:supp:descriptives}a), with 10 countries accounting for 80 percent of all mobility (Fig.~\ref{fig:supp:descriptives}b).
In most cases, countries with a high degree inter-organization mobility also had a high degree of international mobility;
however, the USA and France had a high degree of inter-organization mobility with the corresponding international mobility (Fig.~\ref{fig:supp:descriptives}c).
While the number of publications has increased year-to-year, the mobility and disciplinary makeup of the dataset has not significantly changed across the period of study (Fig.~\ref{fig:supp:pubs_over_time}).
